\chapter[Wind energy applications: Roughness modeling]{Wind energy applications: Roughness modeling}
\label{chapter_6}
%% The following annotation is customary for chapter which have already been
%% published as a paper.
\blfootnote{This chapter was published in the journal Renewable Energy~\cite{Ravishankara_2021}. Other contributing authors were H{\"u}seyin {\"O}zdemir and Edwin van der Weide.}
%% It is only necessary to list the authors if multiple people contributed
%% significantly to the chapter.


\begin{abstract}
 The surface of wind turbine blades are prone to degradation due to exposure to the elements. Rain, hail, insects are among the many causes of turbine blade degradation or erosion. Surface degradation of the wind turbine blades leads to a reduction in the aerodynamic performance, resulting in power losses. The effect of surface degradation is studied by modeling the turbine blade as a rough surface. Surface roughness can be positive (insects or other foreign objects) or negative (erosion, delamination). The individual roughness elements are however very small and it is not always feasible to study the actual degraded surface. Thus various roughness models have been proposed in the literature which eliminate the need to accurately model the degraded surface by representing erosion with a virtual surface and modeling the effect of erosion on the flow quantities near the eroded surface. In this study, the reduction in performance of airfoils due to leading edge roughness is quantified. Different roughness models are investigated and evaluated against theoretical models. Additionally, the effect of roughness on different integral boundary layer quantities like displacement thickness, momentum thickness and skin friction are presented.
\end{abstract}
\input{ch6_roughness/tex/introduction.tex}
%\input{ch6_roughness/tex/num_methods.tex}
\input{ch6_roughness/tex/roughness_theory.tex}
\input{ch6_roughness/tex/airfoil_appln.tex}
\input{ch6_roughness/tex/boundary_layer.tex}
\input{ch6_roughness/tex/conclusions.tex}

\bibliographystyle{dissertation}
\bibliography{ch6_roughness/tex/references}